Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho, and Potential Impacts to Local Chinook Salmon

### 2022-04-13 analysis/paper/paper.Rmd
minor edits in .Rmd that are in submission

---
title: "Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho, and Potential Impacts to Local Chinook Salmon Populations"
author:
  - Michael W. Ackerman:
      email: mike.ackerman@mthoodenvironmental.com
      institute: [mhe_mccall, biomark]
      correspondence: true
  - Nicholas J. Porter:
      email: Nicholas.Porter@merck.com
      institute: [biomark]
      correspondence: false
  - Kevin E. See:
      email: Kevin.See@dfw.wa.gov
      institute: [biomark, wdfw]
      correspondence: false
  - Tara E. Blackman:
      email: tara.blackman@mthoodenvironmental.com
      institute: [mhe_sandy]
      correspondence: false
  - Michael S. Hall:
      email: Mike.Hall1@merck.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richard.Carmichael@merck.com
      institute: [biomark]
      correspondence: false
institute:
  - mhe_mccall: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
  - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
  - wdfw: Washington Department of Fish and Wildlife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
  - mhe_sandy: Mount Hood Environmental, 39085 Pioneer Boulevard \#100 Mezzanine, Sandy, Oregon, 97055, USA 
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: true
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua      
fontsize: 12pt
mainfont: Times New Roman
bibliography: 
- references.bib # Kevin's library
- AckermanLibrary.bib # Mike's library
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
keywords: |
  Northern Pikeminnow; Chinook Salmon; predation; mark-recapture; bioenergetics
---

@@

@@

```{r bio-tbl}
bio_tbl = bio_g_summ %>%
  mutate(`Date Range` = season) %>%
  select(season,
         `Date Range`,
         name,
         cumm_g) %>%
  mutate(name = str_sub(name, 2, 3)) %>%
  rename(Season = season,
         `Percent Fish in Diet` = name,
         `g Fish Consumed` = cumm_g) %>%
  mutate(`Date Range` = recode(`Date Range`,
                               Fall = "Sep 15 - Nov 30",
                               Spring = "Mar 1 - May 31")) %>%
  kable(booktabs = T,
        caption = "Bioenergetics model runs to estimate the grams of fish consumed during the fall DSR and spring NRR Chinook Salmon emigrations. The date ranges for each model run, the assumed proportion of fish in the diet, and the total grams of fish consumed by an individual Northern Pikeminnow to maintain its body weight for a given season are shown. Northern Pikeminnow start and end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
  kable_styling()

@@

```{r gastric-tab}
gastric_tbl %>%
  rename(Survey = event_name,
         Captured = captured,
         Lavaged = lavaged,
         'Stomach Contents' = stmch_contents,
         'Fish Contents' = fish_contents) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Counts of the total number of Northern Pikeminnow captured, gastric lavaged, individuals with stomach contents, and individuals with fish prey contents.") %>%
  kable_styling()

@@

# Figures

@@

```{r adult-imp-p, fig.cap = "Estimated reduction in the annual number of returning adult Chinook Salmon as a result of Northern Pikeminnow predation. Error bars indicate 95% confidence intervals."}
#adult impacts
adult_imp_p = total_impacts %>%
  filter(perc_diet == 60) %>%
  select(p_chnk,
         adult_N,
         adult_Lci,
         adult_Uci) %>%
  mutate(perc_chnk = p_chnk * 100) %>%
  select(-p_chnk) %>%
  ggplot(aes(x = as.factor(perc_chnk),
             y = adult_N)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = adult_Lci,
                    ymax = adult_Uci),
                width = 0) +
  labs(y = "Estimated Adult Chinook",
       x = "Percent Chinook Salmon in Total Fish Consumed") +
  theme(axis.text.x = element_text(size = 11,
                                   color = "black"),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        legend.position = c(.2, .8),
        legend.box = "horizontal")

@@

# Supplementary Materials

>>>>

---
title: "Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho, and Potential Impacts to Local Chinook Salmon"
author:
  - Michael W. Ackerman:
      email: mike.ackerman@mthoodenvironmental.com
      institute: [mhe_mccall, biomark]
      correspondence: true
  - Nicholas J. Porter:
      email: Nicholas.Porter@merck.com
      institute: [biomark]
      correspondence: false
  - Kevin E. See:
      email: Kevin.See@dfw.wa.gov
      institute: [biomark, wdfw]
      correspondence: false
  - Tara E. Blackman:
      email: tara.blackman@mthoodenvironmental.com
      institute: [mhe_sandy]
      correspondence: false
  - Michael S. Hall:
      email: Mike.Hall1@merck.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richard.Carmichael@merck.com
      institute: [biomark]
      correspondence: false
institute:
  - mhe_mccall: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
  - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
  - wdfw: Washington Department of Fish and Wildlife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
  - mhe_sandy: Mount Hood Environmental, 39085 Pioneer Boulevard \#100 Mezzanine, Sandy, Oregon, 97055, USA 
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: true
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua      
fontsize: 12pt
mainfont: Times New Roman
bibliography: 
- references.bib # Kevin's library
- AckermanLibrary.bib # Mike's library
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
keywords: |
  Northern Pikeminnow; Chinook Salmon; predation; mark-recapture; bioenergetics
runninghead: |
  Northern Pikeminnow in Deadwater Slough
---

Running Head: `r rmarkdown::metadata$runninghead`

@@

## Impact Statement

A large population of Northern Pikeminnow in a slow-water reach of the Salmon River, Idaho can potentially consume large numbers of juvenile Chinook Salmon, hindering ongoing recovery efforts.

@@

```{r bio-tbl}
bio_tbl = bio_g_summ %>%
  mutate(`Date Range` = season) %>%
  select(season,
         `Date Range`,
         name,
         cumm_g) %>%
  mutate(name = str_sub(name, 2, 3)) %>%
  rename(Season = season,
         `Fish in Diet (%)` = name,
         `Fish Consumed (g)` = cumm_g) %>%
  mutate(`Date Range` = recode(`Date Range`,
                               Fall = "Sep 15 - Nov 30",
                               Spring = "Mar 1 - May 31")) %>%
  kable(booktabs = T,
        caption = "Bioenergetics model runs to estimate the grams of fish consumed during the fall DSR and spring NRR Chinook Salmon emigrations. The date ranges for each model run, the assumed proportion of fish in the diet, and the total grams of fish consumed by an individual Northern Pikeminnow to maintain its body weight for a given season are shown. Northern Pikeminnow start and end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
  kable_styling()

@@

```{r gastric-tab}
gastric_tbl %>%
  rename(Survey = event_name,
         Captured = captured,
         Lavaged = lavaged,
         'Stomach Contents' = stmch_contents,
         'Fish Contents' = fish_contents) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Counts of the total number of Northern Pikeminnow captured and gastric lavaged. The number of individuals with stomach contents, including those with confirmed fish contents are shown.") %>%
  kable_styling()

@@

# Figure Captions

@@

```{r adult-imp-p, fig.cap = "Estimated “adult equivalents” of the total juvenile Chinook Salmon consumed by Northern Pikeminnow in Deadwater Slough. Error bars indicate 95% confidence intervals."}
#adult impacts
adult_imp_p = total_impacts %>%
  filter(perc_diet == 60) %>%
  select(p_chnk,
         adult_N,
         adult_Lci,
         adult_Uci) %>%
  mutate(perc_chnk = p_chnk * 100) %>%
  select(-p_chnk) %>%
  ggplot(aes(x = as.factor(perc_chnk),
             y = adult_N)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = adult_Lci,
                    ymax = adult_Uci),
                width = 0) +
  labs(y = "Estimated Adult Chinook",
       x = "Percent Chinook Salmon in Total Fish Consumed") +
  theme(axis.text.x = element_text(size = 11,
                                   color = "black"),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        legend.position = c(.2, .8),
        legend.box = "horizontal")

@@

## Supplement - Avian Predation in Deadwater Slough
